\chapter{Study Design}\label{chapter:studydesign}
The main goals of this research were to develop an exergame which can be used for warm up routine before more strenuous physical activity and to evaluate its effectiveness in terms of guiding the user through the process of warming up. In this chapter we outline the research framework, detail the research methods and present the obtained results.
\section{Description of the Experiment}
This section describes the evaluation of the second version of the Immotion exergame. For this purpose, an approach was adopted that uses a mixture of different tools and user study methods. This period of testing of our gamified solution is known as pilot testing. During this period, data has been logged, surveys have been conducted and interviews undertaken. Similarly to the first evaluation, the obtained results are analyzed in order to determine to which level our proposed solution was effective in the context and whether it offered a solution to the problem. Based on the results from the exergame prototype evaluation (Chapter x) and a variety of flow research presented (Chapter x), guidelines were followed which influenced the design and development of the second version of the Immotion exergame. 

\subsection{Introduction and Goals} \label{chapter:goals}
The primary goal of the second study was to investigate whether our exergame solution can be used as an interactive guide for individuals who do not know how to perform warm up routines. In addition, we examined if the exergame can be used as a solution that motivates individuals to warm up before physically more demanding exercises, and provides an enjoyable game experience. \\\\\\Taking this into account, the research questions we address in this study are as follows: 
\begin{enumerate}
\item \textbf{RQ1: Evaluation of effectiveness} - How effective our proposed solution is in guiding the user through the warm up routine compared to the guidance offered by classic (traditional) methods?
%\item Comparing the experience between participants that use the exergame and participants that do not use the exergame. 
%\item Can our proposed solution motivate individuals to undertake warm up procedure?
\item \textbf{RQ2: Evaluation of perceived usefulness and ease of use} - How useful and easy to use our proposed solution is?
\item \textbf{RQ3: Evaluation of the usability} - How usable our proposed solution is? 
\item \textbf{RQ4: Evaluation of the game experience} - How enjoyable and entertaining our proposed solution is? 
\end{enumerate}
%The actual warm up was the original procedure:  
%\begin{itemize}
%\item The warm up routine was derived from sports literature.
%\item The exergame has been designed with this in mind.
%\item A video has been recorded with the original procedure.
%\end{itemize}http://edutechwiki.unige.ch/en/Usability_and_user_experience_surveys
%http://www.nigelbevan.com/papers/What_is_the_difference_between_usability_and_user_experience_evaluation_methods.pdf
In order to evaluate the effectiveness, perceived user experience, usefulness and usability of our gamified solution in the given context, the user base is divided into two groups: \textit{experimental group} and \textit{control group}. The first, experimental group, is the one that interacts with the exergame directly. Contrarily, the control group is presented with the video of a coach (professional) who guides the participant through the warm up routine. This division allows us to infer the influence of our gamified solution, as well as, to assess the main differences in completing the required activities between the two user groups. 
%\subsubsection{Assumptions}
%\begin{itemize}
%\item The participants will be able to perform the requested movements.
%\item The participants will answer all the questionnaires truthfully.
%\item The software and hardware that is used used will function properly.
%\end{itemize}
\subsubsection{Hypotheses}
%Remember to state these in terms of the independent and dependent variables. If it is not immediately clear why you would have a certain hypothesis (it often follows logically from the introduction of the experiment), then include a brief explanation separate from but following the hypothesis. You do not need to state the null hypothesis.%
Based on the research questions outlined in the previous section, the following hypotheses are established to be tested: 
\begin{enumerate}
\item The exergame itself is sufficient for guiding the player through a proper warm up procedure with correct movements. 
\item After the warm up routine is completed using the exergame, the player reached a significantly higher increase in ROM.
%\item Participants that are interactively guided through the warm up procedure using the exergame have perceived exertion during physical activityexertion during a testperceived user experience in terms of PRC scale, compared to the participants who warm up without the exergame.
\item Participants had a more positive perceived warm up experience when using the exergame compared to the participants not using the exergame.  
%\item Participants can easily follow the exergame and perform all the required movements. MODIFY %felt more motivated to undertake warm up exercises when using the exergame compared to the participants not using the exergame.
%\item The exergame will encourage more 
%behavior change compared to the standard warm up %procedure. - for longer study
\end{enumerate}

%Introduce your experiment, and give the reader the specific goals you expect it to address. It is common at this stage to give the reader a hint of your hypotheses (if they are not already hinted at in the Introduction).%
\subsubsection{Apparatus}
The experiment was conducted in the laboratory room in DFKI on (add date). Description of the room - measures to be added. The following equipment has been used:
\begin{itemize}
\item Kinect for Xbox One (2.0 2013) motion sensing input devices by Microsoft used for movement detection and controlling the exergame avatar. 
\item PC running the game engine.
\item Projector used to display the game (video) on the wall in front of the participant.
\item Microsoft Band used for gathering heart rate data.
\item Goniometer used for measuring participants' ROM.
\end{itemize}
TODO: describe where and how is the equipment positioned.

\subsection{Methods} 
In this section we outline the methodology adopted for the Immotion exergame evaluation. For this purpose we utilize the traditional (moderated) usability test since it gives direct input on how real users use the system. %First, we measure users' performance. That is, we measure the time users' will spend performing the warm up exercises with the exergame and with the video. The amount of time users' feel warmed up sufficiently. 
%We decide to follow a mixed-methods approach, and by doing so, utilize both qualitative and quantitative data sources in combination. Second, we measure users' satisfaction level. What the user thinks about his interaction with the product (visually appealing) 
%
%This is a detailed description of the experiment that should allow other researchers (familiar with HCI and experimental design in general, but not familiar with your experiment) to replicate your experiment.%
\subsubsection{Participants}
Demographic details of the participants to be added when the study is completed.\\
Total of n = X individuals participated in the study that has been conducted DATE in DFKI. All participants were students from Saarland University. For recruiting participants, posters were distributed in print, and sent through social media and email (Appendix X). Each participant was given X euros for taking part in the study. All of the participants were amateur athletes who engage in some physical activity on average X times per week. For the study we particularly targeted individuals who exercise in gym or fitness centers and often avoid preforming warm up exercises before more strenuous physical activity. All participants were required to report to the laboratory in gym based clothing, preferably shorts and t-shirt, and all of them performed the required tests in the same location using the same equipment. Before the study, each participant signed a consent form (Appendix X). TODO: This should be updated later with real data.
%Describe your participants (e.g., any relevant demographics, if/how they were divided into categories), including total number, and recruiting approach. Indicate if any incentives were used. Comment on the representativeness of your participants relative to the target population, if their representativeness isn’t immediately obvious.%
\subsubsection{Conditions}
First x participant who applied for the experiment have been accepted. These participants were sent a a short questionnaire that needed to be completed before coming to the experiment. Based on the answers given, the participants were assigned to the control or the experiment group (Appendix X).
Each assigned participant took part in a single test session one hour in duration. During this session, all the participants completed a preliminary questionnaire (Appendix X), after which they performed one exercise session. At the end of the session the participants completed a post-test questionnaire (Appendix X). Two conditions were evaluated:
\begin{enumerate}
\item Exercising with the game guiding through the warm up procedure, projected on a wall in front of the participant.
\item Exercising with a video of a professional (coach) guiding through the exact same warm up procedure as induced by the game, projected on a wall in from of the participant.
\item TO DISCUSS: perhaps third condition participants who warm up without video or the exergame?
\end{enumerate}
Depending on the group, each participant performed exercise that represent one of the conditions.
\subsubsection{Control and Experiment Groups}
The participants are assigned to each group based on the previously completed self-reported survey (Appendix X). This survey was sent to each participant and needed to be completed before the experiment. Based on the answers provided, each participant was assigned to either control or experiment group. The survey assessed participants' perceived fitness level and warm up preferences. TO DISCUSS: Decide how will the participant be divided into groups? Options: level of fitness, warm up preferences, gender, age?
\subsubsection{Measures and Metrics}
During the experiment, the following measures were collected from each participants:
\begin{itemize}
\item ROM measured before the warm up routine and after the warm up routine using goniometer. TODO: what kind of goniometer? To specify.
\item Heart rate data measured using Microsoft Band.
\item Distance during the warm up routine measured using Microsof Kinect. 
\item The warm up routine performed by the participant recorded using Canon XXX camera.
\end{itemize}

%If your experiment is comparing multiple different interfaces or interactive systems or techniques, describe each of them. Screen snapshots of interfaces/systems are particularly useful.%
\subsubsection{Tasks}
In order to interact with the gamified system, the participants in the experiment group were required to perform a set of general movements. By performing these movements, the participant controlled the game avatar and, by doing so, avoided obstacles and collected coins. Based on the data collected from the first survey, in order to successfully finish the game, only movements that are, first, detectable with high accuracy using only one Kinect device and, second, simplistic enough to be accomplished easily without no prior exercise knowledge or experience were required. The movements the participant needed to perform included: 
\begin{itemize}
\item right hand movement up,
\item left hand movement up,
\item jump right,
\item jump left,
\item jump up, 
\item star jump, and
\item squat.
\end{itemize}
Participants who were in the control group and did not interact with the gamified system were required to perform the same set of general movements. However, participants in this group had to follow a video that was projected on the wall in front of them. The video was a recording of a professional (coach) who guided the participants through the warm up routine. By following the video, and thus the coach, the participants were required to execute the same movements as the participant in the experiment group who interacted with the exergame.
%Briefly describe what participants were asked to do with the interactive system(s).%
\subsubsection{Design}
TO DISCUSS this section. Write the formal experimental design (e.g., a 2 x 3 mixed factorial design, more specifically a 2 levels of expertise (between subjects) x 3 interfaces (within subjects) design). 
\subsubsection{Procedure}
%Describe the sequence of activities each participant followed. This should document the experiment from a participant’s perspective, from the moment s/he arrives (e.g., a preliminary questionnaire to obtain X information, followed by five tasks with system A, then a 10 min break, followed by the same five tasks with system B, and finally a semi-structured interview to solicit opinion on Y).%
%The activities each participant of the experiment session followed is presented in this section.\\ 
Before the experiment, the lab environment is set up. The Kinect sensor is placed in a correct position and turned on. The PC running the software is started and the projector is enabled. In each session only one participant is present and guided by the researcher. The activities each participant followed are:
\begin{itemize}
\item The participant completes the preliminary survey.
\item The researcher explains the sensors and tools that are required for the experiment, after which the participant puts them on. 
\item After the researcher confirms that the sensors are placed in a correct position, we start recording heart rate data.
\item The researcher measures the participant's ROM before starting the warm up procedure for the following joints: TODO decide which joint will be measured.
\item After the measurements are completed, the participant rests.
\item The researcher gives a general explanation on the benefits of a proper warm up routine before physically more demanding exercise.
\item The participant moves to the spot marked by the researcher.
\item The researcher starts recording the session. 
\item The warm up procedure begins:
\begin{itemize}
\item If this participant is part of the experimental group, the game starts with the start scene where the participant enters his or her name. After 5 seconds, the game proceeds with scenes in which the participant performs specific movements in order to avoid obstacles and collect coins. The duration of the game is not fixed and it is played up to the point when the participant feels warmed up enough. 
\item  In case the participant is part of the control group, the video that displays a coach who instructs the participants which movements need to be performed. As with with the sessions in the experiment group, the duration of the warm up is not fixed and the video is played up to the point when the participant feels warmed up enough.
\end{itemize}
\item After finishing with the warm up routine, the participant takes a rest. During this period the researcher assesses the ROM of the participant. 
\item After taking the ROM measures, the sensors are removed.
\item The participant completes the post-test survey together with the \textit{RPE} (Rated Perceived Exertion) questionnaire. %https://www.verywell.com/rating-of-perceived-exertion-scale-3119445
\end{itemize}

\subsubsection{Independent and Dependent Variables}
Include exactly how you intend to measure each dependent variable. 
%https://books.google.de/books?hl=de&lr=&id=bPhLeMBLEkAC&oi=fnd&pg=PP1&dq=measuring+the+user+experience&ots=R9KdcqS-vF&sig=2-V-FaMOeX24Rn0pG30lWVdvjSo#v=onepage&q=measuring%20the%20user%20experience&f=false
Independent variables are the things you manipulate or control for, such as design's you are testing or the ages of the respondents. 
Dependent variables are the things you measure, such as success rate, number of errors, user satisfaction, completion time, and many more. Have s clear idea what do you manipulate - independent variables, and what do you want to measure - dependent variables. The most interesting is the intersection - is one design results in a higher task success rate than other. 

Independent and dependent variables can be measured by 4 types of data: nominal, ordinal, interval, and ratio. 
Nominal (categorical) data: groups or categories. Mac vs Windows users, male vs female. These are independent variables, that allows you to segment data by these different groups. Nominal data also includes dependent variables like task success, number of users who clicked link A instead B... 
Ordinal: ordered data - imdb. In user studies this comes from self-reported data. User states if someone is good better worse... These are relative rankings. You report it by frequencies: for example 40 percent said it is good. 

\subsection{Problems/Limitations - Threats to Validity}
Describe any problems/limitations encountered that will help other researchers avoid or account for them if they decide to replicate your experiment.
\section{Results}

This section is an objective report on what the numbers show. You should not try to interpret the meaning of the numbers in this section. Some of the things you may do here are: 
report means and standard deviations in neat tables 
indicate the statistics used and levels of significance 
include graphs, plots, histograms, etc that tell a story about the actual figures obtained 
Only critical raw data and summary statistics should be included in the actual report. However, you must keep all your raw data in a separate archival report, should anyone (a reviewer in the case of real scientific reporting) need more detail than is provided in the paper. 
\section{Discussion}
Interpret the results. Although you should still try to be as objective as possible, the discussion section should illuminate your critical thinking about the results. Explain what the statistics mean, account for anomalies, and so on.
\subsection{Interpretation of Results}
Discuss what you believe the results really mean. For example, if you find a significant difference for some effect, what does that mean to the hypothesis? Is the different seen an important one?
\subsection{Relation to other works}
How do the results you’ve obtained relate to other research findings?
\subsection{Impact for practitioners}
As computer scientists, we are particularly concerned with the implications of our findings on practitioners. Should existing interface constructs be designed differently or used in a new context? Do you have suggestions for new designs? How can the findings be generalized?
\subsection{Critical reflection}
Critical reflection is one of the key foundations of science. You should criticize your work (constructively, if possible), indicate possible flaws, mitigating circumstances, the limits to generalization, conditions under which you would expect your findings to be reversed, and so on.
\subsection{Research agenda}
The best experiments suggest new avenues of exploration. In this section, you should reflect and refine your hypotheses, describe new hypotheses, and suggest future research, ie research that you would do if you continued along this path.
\section{Conclusions}
Summarize the report, and speculate on what is to come.
Acknowledgements. This section should give thanks to the major people (supervisors, associates) and organizations (sponsoring agencies, funders) that helped you. For example, I would like to thank Ben Shneiderman, whose report framework was used to build this one.